Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-07-04 WSSSPE2_report/WSSSPE2_review.tex
undoing more changes

The group felt that training in techniques that promote sustainability has a
range of benefits: career paths, educated reviewers, reproducible science, and
more. There is some information on how that has been approached and assessed,
but more is needed. The group felt that this training is undervalued and that it
is important to communicate the return on investment (ROI) -- both individual ROI
(skills that make scientists more effective and more marketable) and funder ROI
(better use of taxpayer funds; research more likely to be reproducible because
sustainable software exists; better trained reviewers).

@@

\begin{itemize}
\item Michael R. Crusoe and C. Titus Brown. Channeling community contributions to
scientific software: A hackathon experience~\cite{wssspe2_crusoe}

\item Marcus Hanwell, Patrick O'Leary, and Bob O'Bara. Sustainable software
ecosystems: Software engineers, domain scientists, and engineers collaborating
for science~\cite{wssspe2_hanwell}

\item W. Christopher Lenhardt, Stanley Ahalt, Matt Jones, J. Aukema, S. Hampton,
S. R. Hespanh, R. Idaszak, and M. Schildhauer. {ISEES-WSSI} Lessons for
sustainable science software from an early career training institute on open
science synthesis~\cite{wssspe2_lenhardt}

\item Jory Schossau and Greg Wilson. Which sustainable software practices do
scientists find most useful?~\cite{wssspe2_schossau}

@@

\item Jordan Adams, Sai Nudurupati, Nicole Gasparini, Daniel Hobley, Eric
Hutton, Gregory Tucker, and Erkan Istanbulluoglu. Landlab: Sustainable software
development in practice ~\cite{wssspe2_adams}

\item Alice Allen and Judy Schmidt. Looking before leaping: Creating a software
registry~\cite{wssspe2_allen}

\item Carl Boettiger, Ted Hart, Scott Chamberlain, and Karthik Ram. Building
software, building community: lessons from the {ROpenSci} project~\cite{wssspe2_boettiger}

\item Yolanda Gil, Eunyoung Moon, and James Howison. No science software is an
island: Collaborative software development needs in
geosciences~\cite{wssspe2_gil}

\item Ted Habermann, Andrew Collette, Steve Vincena, Werner Benger, Jay Jay
Billings, Matt Gerring, Konrad Hinsen, Pierre de Buyl, Mark K\"{o}nnecke, Filipe
Rnc Maia, and Suren Byna. The Hierarchical Data Format ({HDF}): A foundation for
sustainable data and software~\cite{wssspe2_habermann}

\item Eric Hutton, Mark Piper, Irina Overeem, Albert Kettner, and James Syvitski.
Building sustainable software - the {CSDMS} approach~\cite{wssspe2_hutton}

\item James S. Spencer, Nicholas S. Blunt, William A. Vigor, Fionn D. Malone, W.
M. C. Foulkes, James J. Shepherd, and Alex J. W. Thom. The {H}ighly {A}ccurate
{N-DE}terminant ({HANDE}) quantum {Monte Carlo} project: Open-source stochastic
diagonalisation for quantum chemistry~\cite{wssspe2_spencer}
\end{itemize}

>>>>

The group felt that training in techniques that promote sustainability has a
range of benefits: career paths, educated reviewers, reproducible science, and
more. There is some information on how that has been approached and assessed,
but more is needed. The group felt that this training is undervalued and that it
is important to communicate the return on investment (ROI) -- both individual ROI
(skills that make scientists more effective and more marketable) and funder ROI
(better use of taxpayer funds, research more likely to be reproducible because
sustainable software exists, better trained reviewers).

@@

\begin{itemize}
\item Michael R. Crusoe and C. Titus Brown. Channeling Community Contributions to
Scientific Software: A Hackathon Experience~\cite{wssspe2_crusoe}

\item Marcus Hanwell, Patrick O'Leary, and Bob O'Bara. Sustainable Software
Ecosystems: Software Engineers, Domain Scientists, and Engineers Collaborating
for Science~\cite{wssspe2_hanwell}

\item W. Christopher Lenhardt, Stanley Ahalt, Matt Jones, J. Aukema, S. Hampton,
S. R. Hespanh, R. Idaszak, and M. Schildhauer. {ISEES-WSSI} Lessons for
Sustainable Science Software from an Early Career Training Institute on Open
Science Synthesis~\cite{wssspe2_lenhardt}

\item Jory Schossau and Greg Wilson. Which Sustainable Software Practices do
Scientists Find Most Useful?~\cite{wssspe2_schossau}

@@

\item Jordan Adams, Sai Nudurupati, Nicole Gasparini, Daniel Hobley, Eric
Hutton, Gregory Tucker, and Erkan Istanbulluoglu. Landlab: Sustainable Software
Development in Practice~\cite{wssspe2_adams}

\item Alice Allen and Judy Schmidt. Looking before Leaping: Creating a Software
Registry~\cite{wssspe2_allen}

\item Carl Boettiger, Ted Hart, Scott Chamberlain, and Karthik Ram. Building
Software, Building Community: Lessons from the {ROpenSci} Project~\cite{wssspe2_boettiger}

\item Yolanda Gil, Eunyoung Moon, and James Howison. No Science Software Is an
Island: Collaborative Software Development Needs in
Geosciences~\cite{wssspe2_gil}

\item Ted Habermann, Andrew Collette, Steve Vincena, Werner Benger, Jay Jay
Billings, Matt Gerring, Konrad Hinsen, Pierre de Buyl, Mark K\"{o}nnecke, Filipe
Rnc Maia, and Suren Byna. The Hierarchical Data Format ({HDF}): A Foundation for
Sustainable Data and Software~\cite{wssspe2_habermann}

\item Eric Hutton, Mark Piper, Irina Overeem, Albert Kettner, and James Syvitski.
Building Sustainable Software - The {CSDMS} Approach~\cite{wssspe2_hutton}

\item James S. Spencer, Nicholas S. Blunt, William A. Vigor, Fionn D. Malone, W.
M. C. Foulkes, James J. Shepherd, and Alex J. W. Thom. The {H}ighly {A}ccurate
{N-DE}terminant ({HANDE}) Quantum {Monte Carlo} Project: Open-source Stochastic
Diagonalisation for Quantum Chemistry~\cite{wssspe2_spencer}
\end{itemize}
